---
title: "COVID"
description: |
  The impact of the vaccine on the number of people infected.
author:
  - name: Shaohu Chen
    url: https://example.com/norajones
date: 08-31-2021
output:
  distill::distill_article:
    self_contained: false
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      cache=TRUE, 
                      message =FALSE, 
                      warning=FALSE, 
                      fig.align = 'center')
library(tidyverse)
library(readxl)
library(DT)
library(formattable)
library(gt)
options(scipen=200)

data <- read_xls("data/owid-covid-data.xls")
my_colors <- c("#05A4C0", "#85CEDA", "#D2A7D8", "#A67BC5", "#BB1C8B", "#8D266E")
my_theme <- theme(plot.background = element_rect(fill = "grey98", color = "grey20"),
                  panel.background = element_rect(fill = "grey98"),
                  panel.grid.major = element_line(colour = "grey87"),
                  text = element_text(color = "grey20"),
                  plot.title = element_text(size = 22),
                  plot.subtitle = element_text(size = 17),
                  axis.title = element_text(size = 15),
                  axis.text = element_text(size = 15),
                  legend.box.background = element_rect(color = "grey20", fill = "grey98", size = 0.1),
                  legend.box.margin = margin(t = 3, r = 3, b = 3, l = 3),
                  legend.title = element_blank(),
                  legend.text = element_text(size = 15),
                  strip.text = element_text(size=17))
```


```{r}
vaccination <- data %>% 
  select(c(date, total_vaccinations:new_vaccinations_smoothed_per_million, -c(total_boosters_per_hundred,total_boosters))) %>% 
  mutate(date = as.Date(date))

case <- data %>% 
  select(date:new_cases,reproduction_rate) 

```

Figure 1 shows that the vaccine started in December 2020. In April 2021, Canada began to increase the speed of the vaccine, and people began to inject the second vaccine in July. Up to now, almost most people in the vaccinated population have received 2 shots of the vaccine.

```{r Fig1 ,fig.width=15, fig.height= 9}
vaccination_num <- vaccination %>%
    select(date, total_vaccinations, people_vaccinated,people_fully_vaccinated) %>%
    gather(key = group_var, value = "Cases", -date, na.rm = TRUE) %>%
    group_by(date, group_var) %>%
    summarise(n = sum(Cases, na.rm = F), .groups = "drop_last") %>%
    mutate(label = if_else(date == max(date), as.character(group_var), NA_character_)) 

vaccination_num %>% ggplot(aes(x=date, y = n, color=group_var)) + 
    geom_line(size = 1.5) +
#     geom_label_repel(aes(label = label), nudge_x = 2, hjust=0, na.rm = TRUE, label.size = 0.1, size=3, segment.size = 0.1)
    scale_color_manual(values = my_colors) +
    scale_linetype_manual(values=c("solid", "twodash", "dotted")) +
    coord_cartesian(clip = 'off') +
    scale_y_continuous(labels = scales::comma) +
    scale_x_date(date_breaks = "months" , date_labels = "%b-%y") +
    my_theme + theme(axis.title.x = element_blank()) +
    labs(title = "Reported vaccinations in Time", subtitle = "2021", y = "Frequency")
```

Table 1 shows the specific situation of the vaccine injection, and adds a comparison of the number of people who have received two injections of vaccine and one injection. Through searching, it can be found that the number of people who have received two injections from 2021-07-03 has exceeded that of those who have received two injections. The number of people who have received one shot is nearly 10 times the number of people who have received one shot now, and the goal of covering all two shots has basically been achieved.

```{r Tab1}
vaccination_num %>% select(date : n) %>% 
  filter(date > as.Date("2021-01-04")) %>% 
  pivot_wider(names_from = group_var, 
              values_from = n) %>% 
  mutate(people_fully_vaccinated_percentage =(people_fully_vaccinated /
                                                c(people_vaccinated- people_fully_vaccinated))) %>% 
  mutate(people_fully_vaccinated_percentage = 
           percent(people_fully_vaccinated_percentage)) %>% 
    mutate_each(funs(prettyNum(., big.mark=",")))%>%  
  DT::datatable(options = list(pageLength = 4)) 

```

Figure 2 shows the changes in the number of confirmed cases and the infection rate, and added the three time points mentioned in Figure 1. It can be seen that with the start of vaccination, the number of infections has dropped significantly, but because of the emergence of vaccines, people have begun to less pay attention to their own anti-epidemic behavior has caused the number of infected people to usher in a wave of rise. When the second vaccination started in April, the number of confirmed diagnoses also began to decline, but after July, the number of people who wanted to be vaccinated has basically been covered and the virus has mutated, leading to an increase in the number of infections.


```{r Fig2, fig.width=17, fig.height=7}
case %>% pivot_longer(cols = total_cases:reproduction_rate,
                      names_to = "infection",
                      values_to = "count") %>% 
  ggplot(aes(x = as.Date(date), y = count, color = infection )) +
  geom_line(size = 1.5) +
  facet_wrap(~infection,
             scales = "free_y") +
  theme_bw()+
  geom_vline(xintercept = c(as.Date("2020-12-15"),as.Date("2021-04-05"),as.Date("2021-07-15")),
                            color = "darkgreen", 
             linetype="dashed", size = 1) +
  xlab("Date") +
  scale_color_manual(values = my_colors) +
    scale_linetype_manual(values=c("solid", "twodash", "dotted")) +
    coord_cartesian(clip = 'off') +
    scale_y_continuous(labels = scales::comma) +
    my_theme + theme(axis.title.x = element_blank(),legend.position = "bottom") +
  labs(title = "Reported Cases in Time", subtitle = "2021", y = "Frequency")
```

Table 2 shows the changes in the number of vaccine injections and the number of infections, which further confirms that the number of infections fluctuates with the progress of the vaccine. The number of infections decreased in February and May, but the number of infections increased in April and August. 

```{r Tab2}
p1 <- data %>% select(date, new_cases_per_million, new_vaccinations_smoothed_per_million,reproduction_rate) %>%
  mutate(month=format(date,"%Y-%m")) %>%
  group_by(month) %>% 
  summarise(new_cases_per_million = mean(new_cases_per_million, na.rm = T),
            new_vaccinations_smoothed_per_million = mean(new_vaccinations_smoothed_per_million, na.rm = T),
            reproduction_rate = mean(reproduction_rate, na.rm = T)) %>% 
  filter(month > '2020-11') %>% 
  rename("Date" = month)

p2 <- data %>% select(date, new_cases_per_million,reproduction_rate,new_vaccinations_smoothed_per_million) %>%
  mutate(year = format(date, "%Y")) %>%
  filter(year == "2020") %>% 
  group_by(year)%>% 
  summarise(new_cases_per_million = mean(new_cases_per_million, na.rm = T),
            reproduction_rate = mean(reproduction_rate, na.rm = T),
            new_vaccinations_smoothed_per_million = mean(new_vaccinations_smoothed_per_million, na.rm = T)) %>% 
  rename("Date" = year ) %>% 
  full_join(p1) %>% 
   gt() %>%
  tab_header(title = "Changes in reproductive rates and new cases and new vaccinations about COVID-19 in Canada",
             subtitle = "Profile of average million people of 2020-01-26 to now") %>% 
  tab_source_note(md("Source: Our World in Data")) %>% 
  cols_label(
    new_vaccinations_smoothed_per_million = html("New Vaccinations<br>(per million)"),
    new_cases_per_million = html("New Cases<br> (per million)"),
    reproduction_rate = html("Reproduction<br>  Rate")
  )

p2
```

In conclusion, if we want to further control the increase in the number of infected people, we need to increase people's awareness of prevention and control, frequent disinfection in public places, and the widespread popularization of vaccines and the development of vaccines for mutant viruses.
